---
title: "Lab: Lesson Name"
author: "Author Name"
date-modified: today
format:
  html:
    embed-resources: true
---

# 🛑 Formatting guide (delete this section before posting to the LMS) 🛑

## File naming

Name this file using the convention `lab_lesson_name.qmd`. For example, `lab_introduction_to_writing_functions_in_r.qmd` for the Introduction to Writing Functions in R lesson.

## YAML header

The YAML header should always include `title`, `author`, and `date-modified`. The `format` block should always include `embed-resources: true` so that the rendered HTML file is self-contained.

## Header hierarchy

Use the following header levels consistently throughout the lab:

- `# Load packages` — level-one header (appears once, at the top)
- `# Overview` — level-one header (appears once, after Load packages)
- `# Task N` — level-one header for each task (e.g., `# Task 1`, `# Task 2`)
- `## Notes for students` — level-two header, used after code chunks when explanation is needed
- `## Question N` — level-two header for each follow-up question (e.g., `## Question 1`)

There should be one blank line immediately above all headers.

## Inline code and formatting conventions

- All R code being **run** should be placed in R code chunks.
- All R code being **discussed in text** (but not run) should be written inline using backticks. For example: "Run `install.packages("dplyr")` in the R console."
- R **package names** discussed in text should be formatted as inline code (e.g., "the `dplyr` package").
- **Function names** discussed in text should be formatted as inline code with empty parentheses appended (e.g., "the `mutate()` function").
- **Data frame names**, **variable names**, **file names**, and other key terms should be written in **bold** (e.g., "the **clients_may** data frame", "please download **study_data.csv**").

## Spacing around code chunks

- There should be one blank line immediately above each code chunk.
- There should be at least one blank line immediately below each code chunk.
- If a code chunk is immediately followed by a level-one header (`#`), there should be two blank lines below the code chunk.

## Notes for students

Use a `## Notes for students` section after a code chunk when you want to explain *why* the code works the way it does, flag common mistakes, or connect the current concept to something learners will see later. Notes should be written as a numbered list. Each note should be at least one full sentence. Notes are not required after every code chunk — use them when they add genuine value.

## Question types

Questions follow immediately after the task (or code chunk) they assess. Supported question types include:

- **Fill-in-the-blank / numerical response:** State the question, then provide the answer as a bullet (see Question 1 example below).
- **Multiple choice with distractors:** Present 3–4 code chunks, each preceded by a comment explaining whether it is correct or incorrect and why (see Question 2 example below). Incorrect options should reflect plausible mistakes learners might make.

---


# Load packages

Remember, it's considered a best practice to load all the packages that our file will use at the top of the file. If we have not yet installed a package, we will need to do so by running `install.packages("package name")` in the R console. For example, to install the `dplyr` package, we would run `install.packages("dplyr")`. However, we typically **do not** want to type the code to install the package here in our Quarto file because we only need to install the package once on a given computer. Not every time we run our R code.

```{r}
#| label: load-packages
library(dplyr, warn.conflicts = FALSE) # The "warn.conflicts" part is optional
```


# Overview

The Epi-Workbench Research Consortium (EWBRC) is a fictional public health research organization used throughout Epi-Workbench labs. Scenarios are set within EWBRC to give learners a consistent institutional context across labs.

Provide a brief narrative overview of the lab scenario here. Describe the public health problem or research question EWBRC is working on, the dataset being used, and what learners will accomplish by the end of the lab. Two to four sentences is typically sufficient.


# Task 1

Task 1 typically asks learners to simulate data using `tribble()` or another inline method, or to download an external data file. If downloading, provide clearly labeled links:

* [data_file_1.csv](link)
* [data_file_2.csv](link)

```{r}
df <- tribble(
  ~var1, ~var2, ~var3,
  "a",   1,     TRUE,
  "b",   2,     FALSE
)
```

## Notes for students

1. Explain something important about the code above that a learner might not immediately understand. What does it do? Why was this approach chosen over an alternative?

2. Add a second note if there is another concept worth highlighting. If there is nothing substantive to add, omit the Notes for students section entirely.


# Task 2

Task 2 often asks learners to inspect the structure of the data they just created or imported. Use the `str()` or `glimpse()` functions.

```{r}
glimpse(df)
```

## Question 1

*Example of a fill-in-the-blank / numerical response question.*

The **df** data frame contains ____ rows and ____ columns.

* The **df** data frame contains 2 rows and 3 columns.


# Task 3

Subsequent tasks ask learners to perform data management steps, run analyses, or create visualizations — depending on the topic of the lab. Each task should require learners to write and run code. Tasks should build on one another sequentially.

```{r}
#| eval: false

# Example code chunk — replace with task-specific code.
df |>
  mutate(new_var = var2 * 2)
```

## Question 2

*Example of a multiple-choice question with distractors.*

Which of the following code chunks will correctly create a new column called **new_var** that doubles the values in **var2**?

```{r}
#| eval: false

# Incorrect: This code overwrites df with a vector instead of a modified data frame.
df <- df$var2 * 2
```

```{r}
#| eval: false

# Incorrect: The mutate() function requires a data frame as input. Passing df$var2
# directly discards the rest of the data frame.
mutate(df$var2, new_var = var2 * 2)
```

```{r}
#| eval: false

# Correct.
df <- df |>
  mutate(new_var = var2 * 2)
```

```{r}
#| eval: false

# Correct. This is equivalent to the pipe-based version above.
df <- mutate(df, new_var = var2 * 2)
```


# Task 4

Continue adding tasks as needed. There is no fixed number of tasks required, but labs typically include between 6 and 14 tasks. Each task should be focused and completable in an hour.

```{r}
# Task code here
```

## Notes for students

1. Add notes when helpful. Omit this section when there is nothing substantive to add.

## Question 3

Add a follow-up question here. At least one question per task is typical, though not every task requires a question.

* Correct answer here.
