---
title: "NTRHD Women’s Health Study"
output: html_notebook
---

# Overview

The North Texas Regional Health Department (NTRHD) is conducting a women's health study in Tarrant County. Data for this study is collected remotely at study sites around Tarrant County, scanned, and sent back to the data processing center to be processed. The first batch of survey forms just arrived this morning and needs to be processed.

```{r}
library(dplyr)
```

# Task 1

Please click the links below to download the study forms to your computer.  Please skim through the study forms to get a feel for their layout and the kind of information they contain.

* [Participant 1](https://www.dropbox.com/s/r41fxfiw85qn7ce/Cooper%2C%20Samantha.pdf?dl=1)

* [Participant 2](https://www.dropbox.com/s/8jc4xtakcum0o2m/Rodriguez%2C%20Leslie.pdf?dl=1)

* [Participant 3](https://www.dropbox.com/s/19hf3iqx5krgm3b/Smith%2C%20Jane.pdf?dl=1)


# Task 2

After you have skimmed through the study forms, please create an R data frame named `form_20` that contains information about the 3 study participants listed above in an R Notebook file. The data frame you create should contain the following information:

* The date the form was received   
* The contact type   
* The participant's first name   
* The participant's last name   
* The participant's social security number   

Please make sure you give each variable a meaningfully descriptive name. For example, use something like `name_first` rather than `var1` or `x1`.

## Question:

1. Which of the following code chunks would correctly create the `form_20` data frame and adheres to the coding best practices described in R4Epi?

```{r eval=FALSE}
# Incorrect: The data frame is named form.20 instead of form_20.
form.20 <- data.frame(
  date_received = rep("2013-08-22", 3),
  contact_type  = c("mail", "visit", NA),
  name_first    = c("Samantha", "Leslie", "Jane"),
  name_last     = c("Cooper", "Rodriguez", "Smith"),
  ssn           = c(NA, "321-54-9876", "496-86-5327")
)
```

```{r eval=FALSE}
# Incorrect: Uses equal sign instead of the assignment operator.
form_20 = data.frame(
  date_received = rep("2013-08-22", 3),
  contact_type  = c("mail", "visit", NA),
  name_first    = c("Samantha", "Leslie", "Jane"),
  name_last     = c("Cooper", "Rodriguez", "Smith"),
  ssn           = c(NA, "321-54-9876", "496-86-5327")
)
```

```{r}
# Correct answer
form_20 <- data.frame(
  date_received = rep("2013-08-22", 3),
  contact_type  = c("mail", "visit", NA),
  name_first    = c("Samantha", "Leslie", "Jane"),
  name_last     = c("Cooper", "Rodriguez", "Smith"),
  ssn           = c(NA, "321-54-9876", "496-86-5327")
)
```

```{r eval=FALSE}
# Incorrect: Variable names are not descriptive.
form_20 <- data.frame(
  date   = rep("2013-08-22", 3),
  type   = c("mail", "visit", NA),
  name   = c("Samantha", "Leslie", "Jane"),
  name_2 = c("Cooper", "Rodriguez", "Smith"),
  num    = c(NA, "321-54-9876", "496-86-5327")
)
```

```{r eval=FALSE}
# Incorrect: No space around the assignment operators and commas
form_20<-data.frame(
  date_received=rep("2013-08-22",3),
  contact_type=c("mail","visit",NA),
  name_first=c("Samantha","Leslie","Jane"),
  name_last=c("Cooper","Rodriguez","Smith"),
  ssn=c(NA,"321-54-9876","496-86-5327")
)
```

```{r eval=FALSE}
# Incorrect: No quotation marks around the date received
form_20 <- data.frame(
  date_received = rep(2013-08-22, 3),
  contact_type  = c("mail", "visit", NA),
  name_first    = c("Samantha", "Leslie", "Jane"),
  name_last     = c("Cooper", "Rodriguez", "Smith"),
  ssn           = c(NA, "321-54-9876", "496-86-5327")
)
```

```{r eval=FALSE}
# Incorrect: There should not be quotes around the NA values that represent missing.
form_20 <- data.frame(
  date_received = rep("2013-08-22", 3),
  contact_type  = c("mail", "visit", "NA"),
  name_first    = c("Samantha", "Leslie", "Jane"),
  name_last     = c("Cooper", "Rodriguez", "Smith"),
  ssn           = c("NA", "321-54-9876", "496-86-5327")
)
```

## Notes for students

1. I don't expect you to have this notes section in your R Notebook.   

2. Notice that I manually added space between the variable names and the equals sign. I did this so that they all line-up. You don't have to do this, but I think it makes the code easier to read.

3. Notice that I used the `rep()` function to enter the values for `date_received`. I did this because the value was the same for all three participants -- "2013-08-22". The `rep()` function creates a vector. The first argument to the `rep()` function determines the value that the vector contains. The second argument to the `rep()` function determines the number of times that value will be repeated. So, c("2013-08-22", "2013-08-22", "2013-08-22") and rep("2013-08-22", 3) mean the same thing to R.

4. Notice that the date here is just a character string to R. It doesn't have any true date value. In other words, "2013-08-22" isn't 1 day before "2013-08-23" to R. They are both just character strings like "Jane" and "Samantha". Later, we will learn how to tell R that these are actually dates.

5. Notice that Jane Smith is missing information about contact type and Samantha Cooper's social security number is missing. In R, we represent missing data with an `NA`. Make sure you capitalize `NA` and don't use any spaces or quotation marks. Also, make sure you use `NA` instead of writing `"Missing"` or something like that. 

  * By default, R considers `NA` to be a logical-type value (as opposed to character or numeric). for example:

```{r}
typeof(NA)
```

  * However, you can tell R to make `NA` a different type by using one of the more specific forms of `NA`. For example:

```{r}
typeof(NA_character_)
```

```{r}
typeof(NA_integer_)
```

```{r}
typeof(NA_real_)
```

  * Most of the time, you won't have to worry about this because R will take care of converting `NA` for you. What do I mean by that? Well, remember that every vector can have only one type. So, when you add an `NA` (logical by default) to a vector with character values as we did above `c("mail", "visit", NA)`, that would cause you to have two character values and one logical value in the same vector, which is not allowed. Therefore, R will automatically convert the `NA` to `NA_character_` for you behind the scenes. This is a concept known as "type coercion" and you can read more about it [here](https://r4ds.had.co.nz/vectors.html#coercion) if you are interested. As I said, most of the time you don't have to worry about type coercion -- it will happen automatically. But, sometimes it doesn't and it will cause R to give you an error. I mostly encounter this when using the `if_else()` and `case_when()` functions, which we will discuss later.

6. Notice that I wrote "name_last" and "name_first" instead of "last_name" and "first_name". This was intentional. When you have multiple related variables, it's good practice to start their variable names with the same word. Addresses are another common example (address_street, address_city, and address_state). It makes these related variables easier to find and work with in the future if we need to do something with all of them at once. All we have to do is sort our variable names alphabetically, and all the variables will be grouped together.

7. Notice that I named the variable representing each participant's social security number `ssn`. Based on the style guidelines I gave you, we could have a reasonable discussion about whether or not this name qualifies as sufficiently informative. In many cases, a three-letter acronym would not be sufficiently informative; however, I would argue that in this case it is based on the context of the data (i.e., this is all sociodemographic data) and based on the fact that "ssn" is a really commonly used acronym. Another good example is `dob` to represent date of birth. 

8. Notice that I created the social security numbers as character strings. Why? Well, for starters, they include dashes, which are not allowed in numeric vectors. But, I could have just written them without the dashes (e.g., `321549876`). So, why didn't I? There is a little bit of gray area here, but when deciding whether to make numbers character strings or not I tend to ask myself something like, "would it make sense to subtract (or add or whatever) two of these values?" For example, I could subtract `321549876` from `496865327` (the two social security numbers from above) to get 175315451. But, does 175315451 mean anything? No. Therefore, the ssn's are "numbers", but they aren't really "numeric data." Other good examples of this are phone numbers and participant ID numbers. 

# Task 3

After you create the `form_20` data frame, please view it in the Notebook to make sure your program worked as expected. 

## Question:

2. Which of the following code chunks could you use to view the `form_20` data frame?

```{r}
# Correct
print(form_20)
```

```{r}
# Correct
form_20
```

```{r eval=FALSE}
# Incorrect: This code will return Error: unexpected symbol in "print form_20"
print form_20
```

```{r eval=FALSE}
# Incorrect: This code will return Error in view(form_20) : could not find function "view"
view(form_20)
```

# Task 4

Additionally, please view the structure of the `form_20` data frame in your R Notebook using either the `str()` function or the `dplyr::glimpse()` function.

```{r}
# Using the base R str() function
str(form_20)
```

```{r}
# Using dplyr's glimpse() function
# You would normally add library(dplyr) to the top of your R Notebook and then
# omit the "dplyr::" before the "glimpse(form_20)". I did it below to make it 
# clear that the glimpse() function is from the dplyr package. 
dplyr::glimpse(form_20)
```

## Question:

3. The `form_20` data frame contains ____ rows and ____ columns.

* The `form_20` data frame contains 3 rows and 5 columns.

## Notes for students

1. I don't expect you to have this notes section in your R Notebook.   

2. Why did I have you print the structure of the data? In this case, it isn't all that useful. Our data frame is so small that just viewing it directly makes a lot of sense. However, I wanted to introduce you to the `str()` and `glimpse()` functions because they can be useful when you have much larger data frames. Which one you use is really just a matter of preference.

# Optional feedback

In your R Notebook, please feel free to leave any comments about the usefulness of this lab. Which parts were helpful? What could I do to improve it? What is still unclear?