---
title: "Culminating Exercise: Column-wise Operations with dplyr"
author: "[Your Name]"
date-modified: [Enter Date]
format:
  html:
    embed-resources: true
---


# Load packages

Remember, it's considered a best practice to load all the packages that our file will use at the top of the file. If we have not yet installed a package, we will need to do so by running `install.packages("package name")` in the R console. For example, to install the `dplyr` package, we would run `install.packages("dplyr")`. However, we typically **do not** want to type the code to install the package here in our Quarto file because we only need to install the package once on a given computer. Not every time we run our R code.

```{r}
library(dplyr, warn.conflicts = FALSE)
library(tibble, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(stringr, warn.conflicts = FALSE)
library(snakecase, warn.conflicts = FALSE)
```


# Overview

You are a data analyst on the chronic disease surveillance team at the Epi-Workbench Research Consortium, supporting a multi-site study tracking cardiovascular disease risk factors among adults enrolled at community health centers across six urban regions. Each quarter, the participating centers submit a wide-format data extract containing self-reported behavioral risk factor scores, clinical biometric measurements, and binary screening result flags for conditions such as hypertension, prediabetes, dyslipidemia, and metabolic syndrome. This quarter's extract has arrived in the familiar state of real-world program data: column names are inconsistent across centers, a sentinel code of `–9` has been used in several behavioral score fields to denote "not collected," and the screening result columns follow no shared naming convention. Your task is to clean, standardize, and summarize the dataset so that a quarterly risk profile report can be delivered to the EWBRC's chronic disease advisory board. Along the way, you will apply `dplyr`'s column-wise operation tools — `across()`, `if_all()`, `if_any()`, and `rename_with()` — to perform these cleaning and summary tasks efficiently across all relevant columns at once, in a way that will scale gracefully as additional sites and risk factor variables are added to the study in future quarters.


# Task 1

Run the code below to create the simulated **cvd_risk_wide** dataset.

```{r}
set.seed(101)

cvd_risk_wide <- tibble(
  center_id = c(
    rep("Center_1", 8), rep("Center_2", 8), rep("Center_3", 7),
    rep("Center_4", 7), rep("Center_5", 6), rep("Center_6", 6)
  ),
  age_group = sample(
    c("18-34", "35-49", "50-64", "65+"), 42, replace = TRUE
  ),
  sex = sample(c("Male", "Female"), 42, replace = TRUE),

  # Behavioral risk factor columns with prefix beh_
  beh_diet = as.integer(c(
    2, 3, 3, 1, 0, 4, NA, 2,
    3, 1, 4, 0, 2, 4, 1, 3,
    0, 3, 2, 1, 3, 4, 2,
    NA, 1, 3, 0, 2, 4, 1,
    3, 2, -9, 0, 1, 4,
    2, 0, 3, 1, -9, 2
  )),
  beh_activity = as.integer(c(
    1, 3, -9, 2, 4, 0, 1, NA,
    1, 2, 1, 3, 0, 2, 4, 1,
    3, 0, 1, 2, 1, NA, 3,
    2, 0, 1, 3, 0, 2, 4,
    1, -9, 2, 3, 0, 1,
    4, 2, -9, 1, 3, 0
  )),
  beh_smoking = as.integer(c(
    0, 1, 2, -9, 3, 1, 0, 2,
    1, -9, 3, 2, 0, 1, NA, 2,
    3, 1, 0, 3, 2, 1, 0,
    1, 2, 1, 3, 0, 1, 2,
    NA, 1, 0, 2, 0, 3,
    1, 2, 0, -9, 1, 3
  )),
  beh_alcohol = as.integer(c(
    3, 1, 0, 2, -9, 1, 2, 0,
    2, 3, 1, -9, 0, 2, 1, 4,
    1, 2, 3, -9, 0, 2, NA,
    3, 1, 2, 0, 1, -9, 3,
    2, 0, 1, 3, 2, -9,
    0, 3, 1, 2, 4, -9
  )),
  beh_sleep = as.integer(c(
    2, 0, 1, 3, 2, -9, 1, 0,
    -9, 2, 3, 1, NA, 0, 2, 1,
    2, 3, 1, 0, -9, 1, 2,
    0, 3, 2, 1, 4, 0, 1,
    1, 3, 2, 4, 0, 1,
    1, 1, 2, 0, 3, 1
  )),
  beh_stress = as.integer(c(
    1, 2, 3, 0, 1, 2, 1, 3,
    0, 1, 2, 3, 1, 4, 0, 2,
    1, 2, 0, 3, 1, 2, -9,
    1, 2, 0, -9, 3, 2, 1,
    0, 2, 3, 1, 2, 0,
    3, 0, 1, 2, 0, 1
  )),

  # Screening result columns with inconsistent names
  scrHBP = as.integer(c(
    1, 0, 0, 1, 0, 0, 1, 0,
    0, 1, 0, 0, 1, 0, 0, 1,
    0, 0, 1, 0, 0, 1, 0,
    1, 0, 0, 1, 0, 0, 0,
    1, 0, 0, 1, 0, 0,
    0, 1, 0, 0, 1, 0
  )),
  scr_PreDM = as.integer(c(
    0, 1, 0, 0, 1, 0, 0, 0,
    1, 0, 0, 1, 0, 0, 1, 0,
    0, 1, 0, 0, 1, 0, 0,
    0, 1, 0, 0, 1, 0, 0,
    0, 1, 0, 0, 1, 0,
    1, 0, 0, 1, 0, 0
  )),
  Scr_dyslipid = as.integer(c(
    0, 0, 1, 0, 0, 1, 0, 0,
    0, 0, 1, 0, 0, 1, 0, 0,
    1, 0, 0, 0, 0, 0, 1,
    0, 0, 1, 0, 0, 0, 1,
    0, 0, 1, 0, 0, 0,
    0, 0, 1, 0, 0, 1
  )),
  METSYN = as.integer(c(
    0, 0, 0, 0, 1, 0, 0, 1,
    0, 0, 0, 0, 0, 0, 0, 1,
    0, 0, 0, 1, 0, 0, 0,
    0, 0, 0, 1, 0, 0, 0,
    0, 0, 0, 0, 1, 0,
    0, 0, 0, 0, 0, 1
  )),
  scrObesity_1 = as.integer(c(
    0, 0, 0, 0, 0, 0, 1, 0,
    0, 0, 0, 0, 0, 1, 0, 0,
    0, 0, 0, 0, 1, 0, 0,
    0, 0, 0, 0, 0, 1, 0,
    0, 0, 1, 0, 0, 0,
    0, 0, 0, 0, 0, 1
  ))
)

cvd_risk_wide
```


# Task 2

Inspect the structure of **cvd_risk_wide** and determine how many integer columns it contains.

```{r}

```

## Question 1

How many numeric (integer) columns does **cvd_risk_wide** contain?


# Task 3

Standardize the column names in **cvd_risk_wide**.

a. Use `rename_with()` with `to_snake_case()` to convert every column name to snake_case.
b. Use `rename()` to add the `scr_` prefix to **metsyn**.
c. Store the result as **cvd_clean_names** and display the column names.

```{r}

```

## Question 2

After running the Task 3 code, which of the following is a column name in **cvd_clean_names**?

See our course LMS for the answer choices and submit your response.

# Task 4

Use `mutate()` with `across()` to replace the sentinel value `–9` with `NA` in all columns whose names start with `"beh_"`. Use the `.names` argument set to `"{.col}_clean"` so that the recoded values appear in new columns alongside the originals. Store the result as **cvd_recoded**, then display the original and cleaned behavioral columns side by side.

```{r}

```

## Question 3

How many sentinel values (`–9`) were replaced with `NA` across all six behavioral columns?


# Task 5

Rerun your sentinel recode from Task 4, but this time overwrite the original behavioral columns instead of creating new `_clean` columns. Store the result as **cvd_recoded**.

```{r}

```

## Question 4

The **cvd_recoded** data frame contains ____ rows and ____ columns.


# Task 6

Use `filter()` with `if_all()` to retain only the rows in **cvd_recoded** that have no missing values across all behavioral columns. Store the result as **cvd_complete**, then calculate the number of rows excluded.

```{r}

```

## Question 5

How many rows were excluded by the `if_all()` complete-case filter in Task 6?


# Task 7

Use `summarise()` with `across()` to compute the count of `NA` values, the mean, and the standard deviation for every behavioral column in **cvd_complete**. Name the output columns using the `{.col}-{.fn}` pattern and store the result as **behavior_summary**. Then reshape **behavior_summary** into long format.

```{r}

```

```{r}

```

## Question 6

In **behavior_summary**, what is the standard deviation of **beh_sleep**, rounded to two decimal places?


# Task 8

Group **cvd_complete** by **center_id** and use `summarise()` with `across()` to compute the mean and median for all behavioral columns. Use the `{.col}-{.fn}` naming pattern and store the wide result as **center_summary_wide**. Then reshape it into long format and store as **center_summary_long**.

```{r}

```

```{r}

```

## Question 7

Which center had the highest mean **beh_diet** score?


# Task 9

Produce the same center-level mean and median summary, but this time use a reshape-then-summarize approach. Start with **cvd_complete**, use `pivot_longer()` to create columns named **behavior** and **score**, then group and summarise. Store the result as **center_summary_long**.

```{r}

```

## Question 8

Which center had the lowest median **beh_stress** score?


# Task 10

Use `filter()` with `if_any()` to identify participants in **cvd_complete** who had at least one positive screening result across all screening columns. Store the result as **any_positive** and report the number of rows.

```{r}

```

## Question 9

How many participants in **cvd_complete** had at least one positive result across all screening columns?


# Task 11

Use `mutate()` with `across()` and `where(is.numeric)` to create z-score–normalized versions of all numeric columns in **cvd_complete**. Use the `.names` argument so that each new column name ends with `_z`. Store the result as **cvd_z**.

```{r}

```

```{r}

```

## Question 10

In **cvd_z**, what is the value in the tenth row for **beh_activity_z**, rounded to two decimal places?
