---
title: "Repeated Operations I"
output: html_notebook
---

# Overview

This week, we will practice [writing functions](https://www.r4epi.com/writing-functions.html) and performing [column-wise repeated operations with dplyr](https://www.r4epi.com/column-wise-operations-in-dplyr.html) verbs. In some cases, we will be completing tasks that we already completed in previous modules, but we will do so in a more efficient way. We will use the following data sets for all the tasks in this lab:

* [Click here to download Class Survey 5141.csv](https://www.dropbox.com/s/70q553h37cacjka/Class%20Survey%205141.csv?dl=1). This is the class survey data we used in the conditional operations lab. Feel free to reuse your code from that lab for the tasks below that are repeated from that lab.

* [Click here to view the codebook for the class survey data](https://www.dropbox.com/s/kjyfq8cfur0hzdp/5313%20-%205314%20Day%20One%20Survey%20Codebook.pdf?dl=0).

```{r}
library(dplyr, warn.conflicts = FALSE)
library(readr)
library(stringr)
library(tidyr)
```

# Task 1

Please download `Class Survey 5141.csv` and import it into your R global environment as `class_survey`.

```{r}
class_survey <- read_rds("/Users/bradcannell/Dropbox/Datasets/Class Survey/class_survey.Rds")
```

# Task 2

Please look at the codebook for the `class_survey` data. Several questions on the survey included "Not Sure" as a response option. They were questions 2, 4, 6, 7, 9, 10, 13, 14, 15, 16, 17, 18, 20, and 26. Please rename each of the columns contains these questions to `required`, `still_enroll`, `package`, `genhealth`, `asthma`, `seatbelt`, `insurance`, `persdoc`, `hispanic`, `race`, `veteran`, `marital`, `employment`, `sex`, and `color`.

```{r}
class_survey |> 
  rename(
    required     = q2,
    still_enroll = q4,
    package      = q6,
    genhealth    = q7,
    asthma       = q9,
    seatbelt     = q10,
    insurance    = q13,
    persdoc      = q14, 
    hispanic     = q15,
    race         = q16,
    veteran      = q17, 
    marital      = q18, 
    employment   = q20, 
    sex          = q26,
    color        = q27
  )
```

## Notes for students

1. In the code above, I decided to give a more informative name to each of the columns that we will analyze in the next few tasks. I would have probably had to keep looking at the codebook throughout the analysis if I didn't rename them. I find doing that frustrating, and the time spent renaming the columns well worth it.

2. There isn’t really “repetition” in the code above, per se, but there is still a way that we can make the code easier to update in the future. Start by creating a character vector containing the names of the columns we want to operate on.

```{r}
cols_9 <- c(
  "required", "still_enroll", "package", "genhealth", "asthma", "seatbelt", 
  "insurance", "persdoc", "hispanic", "race", "veteran", "marital", "employment", 
  "sex", "color"
)
```

## Notes for students

1. In the code above, I created a character vector containing the names of all the columns that have a “Not Sure” response option. In the code chunk below, I will use this character vector to help rename our columns. The order of the column names in this vector is important.

2. I will also use this character vector of column names to perform the other tasks that follow. There are a couple of nice things about using this strategy. First, it reduces typing and likelihood of making errors. Second, if we all of the sudden realized that we needed to add another column to each of the tasks we are going to perform below, we will only need to update this character vector once instead of each code chunk below separately.

```{r}
class_survey <- class_survey |> 
  rename_with(
    .fn   = ~ cols_9,
    .cols = c(q2, q4, q6, q7, q9, q10, q13:q18, q20, q26:q27)
  )
```

## Notes for students

1. We’ve used `rename_with()` before to apply the `str_to_lower()` function to all of our column names. However, we can pass any function to the `.fn` argument of `rename_with()`. Passing the character vector, we created earlier as an anonymous function (I wrote the anonymous function above using a purrr-style lambda) to the `.fn` argument has the effect of assigning the names in the character vector to the columns of `class_survey`. It’s very important that (1) the number column names in the character vector matches the number of columns you pass to the `.cols` argument, and (2) the order of the column names in the character vector matches the order of the columns you pass to the `.cols` argument.

## Question 1

Which function can we use to apply other functions (e.g., `str_to_lower()`) to some, or all, of our column names simultaneously in order to change them in some way?

* We can use the `rename_with()` function to apply other functions (e.g., `str_to_lower()`) to some, or all, of our column names _simultaneously_ in order to change them in some way. We can apply functions to individual column names inside of `rename()` and `select()`, but we can't apply those functions to _multiple_ column names simultaneously inside of `rename()` and `select()`. 

# Task 3

Please coerce all of the columns with a Not Sure option from character type to numeric type. Please use the `across()` function to help you remove some of the repetition from your code.

```{r}
# Old way
class_survey |> 
  mutate(
    required     = as.numeric(required), 
    still_enroll = as.numeric(still_enroll), 
    package      = as.numeric(package),
    genhealth    = as.numeric(genhealth),
    asthma       = as.numeric(asthma),
    seatbelt     = as.numeric(seatbelt),
    insurance    = as.numeric(insurance),
    persdoc      = as.numeric(persdoc), 
    hispanic     = as.numeric(hispanic),
    race         = as.numeric(race),
    veteran      = as.numeric(veteran), 
    marital      = as.numeric(marital), 
    employment   = as.numeric(employment), 
    sex          = as.numeric(sex),
    color        = as.numeric(color)
  )
```

```{r}
# With across()
class_survey |> 
  mutate(
    across(
      .cols = c(
        required, still_enroll, package, genhealth, asthma, seatbelt, 
        insurance, persdoc, hispanic, race, veteran, marital, employment, 
        sex, color
      ),
      .fns  = as.numeric
    )
  )
```

```{r}
# With across() and our character vector
class_survey <- class_survey |> 
  mutate(
    across(
      .cols = all_of(cols_9),
      .fns  = as.numeric 
    )
  )
```

## Notes for students

1. Using `across()` removes all of the repetition from our code, which is great. Passing the character vector of column names we previously created to the `.cols` argument (using the `all_of()` tidy-select function) make our code even easier to read and maintain.

# Task 4

Please calculate the number of participants who selected Not Sure as their response for each variable in the class_survey data that included Not Sure as an option. As you complete this task, please write a function that will help remove some of the repetition from your code.

```{r}
# Here's how we might have completed this task prior to this module.
# So much repetition!
class_survey |> 
  summarise(
    required_9     = sum(required == 9, na.rm = TRUE),
    still_enroll_9 = sum(still_enroll == 9, na.rm = TRUE),
    package_9      = sum(package == 9, na.rm = TRUE),
    genhealth_9    = sum(genhealth == 9, na.rm = TRUE),
    asthma_9       = sum(asthma == 9, na.rm = TRUE),
    seatbelt_9     = sum(seatbelt == 9, na.rm = TRUE),
    insurance_9    = sum(insurance == 9, na.rm = TRUE),
    persdoc_9      = sum(persdoc == 9, na.rm = TRUE), 
    hispanic_9     = sum(hispanic == 9, na.rm = TRUE),
    race_9         = sum(race == 9, na.rm = TRUE),
    veteran_9      = sum(veteran == 9, na.rm = TRUE), 
    marital_9      = sum(marital == 9, na.rm = TRUE), 
    employment_9   = sum(employment == 9, na.rm = TRUE), 
    sex_9          = sum(sex == 9, na.rm = TRUE),
    color_9        = sum(color == 99, na.rm = TRUE) # Notice that color uses 99
  )
```

```{r}
# Now, let's write a function.
n_nines <- function(x) {
  sum(x == 9 | x == 99, na.rm = TRUE)
}
```

```{r}
# Use the n_nines function to reduce the amount of repetitive code.
class_survey |> 
  summarise(
    required_9     = n_nines(required),
    still_enroll_9 = n_nines(still_enroll),
    package_9      = n_nines(package),
    genhealth_9    = n_nines(genhealth),
    asthma_9       = n_nines(asthma),
    seatbelt_9     = n_nines(seatbelt),
    insurance_9    = n_nines(insurance),
    persdoc_9      = n_nines(persdoc), 
    hispanic_9     = n_nines(hispanic),
    race_9         = n_nines(race),
    veteran_9      = n_nines(veteran), 
    marital_9      = n_nines(marital), 
    employment_9   = n_nines(employment), 
    sex_9          = n_nines(sex),
    color_9        = n_nines(color) 
  )
```

# Task 5

Please use the across() function to apply the function you created in Task 4 to calculate the number of participants who selected Not Sure as their response for each variable in the class_survey data that included Not Sure as an option. 

```{r}
# Use summarise across to remove even more repetitive code.
class_survey |> 
  summarise(
    across(
      .cols = all_of(cols_9),
      .fns  = n_nines
    )
  )
```

# Task 6 

Please repeat Task 5. This time, however, please pass the code you wrote in your function to the .fns argument as an anonymous function.

```{r}
# Use an anonymous function.
class_survey |> 
  summarise(
    across(
      .cols = all_of(cols_9),
      .fns  = function(x) { 
        sum(x == 9 | x == 99, na.rm = TRUE)
      }
    )
  )
```

# Task 7

Please repeat Task 6. This time, however, please pass your anonymous function to the .fns argument using the purrr-style lambda syntax.

```{r}
# Use the dplyr purr_style lambda syntax.
class_survey |> 
  summarise(
    across(
      .cols = all_of(cols_9),
      .fns  = ~ sum(.x == 9 | .x == 99, na.rm = TRUE)
    )
  )
```

## Question 2

Tasks 4 through 7 above essentially required you to do the same thing in four different slightly different ways. For this particular task, which method did you prefer, and why?

# Task 8

Please repeat Task 7. This time, however, please change the .names argument so that the column names in our results data frame all end with “n_not_sure”. Please assign the results data from the name not_sure_count.

```{r}
# Make the column names more informative.
not_sure_count <- class_survey |> 
  summarise(
    across(
      .cols  = all_of(cols_9),
      .fns   = ~ sum(.x == 9 | .x == 99, na.rm = TRUE),
      .names = "{col}-n_not_sure" # Using a hyphen here makes the next task easier
    )
  )
```

# Task 9

Please restructure not_sure_count so that it has two columns. The first column should contain each of the column names from class_survey that we analyzed in tasks 4 through 8. The second column of not_sure_count should contain the number of Not Sure responses for each of the column names from class_survey that we analyzed in tasks 4 through 8.

```{r rows.print = 15}
# When there are lots of columns to look through, pivot longer makes it easier to
# view the results.
not_sure_count |> 
  pivot_longer(
    cols          = everything(),
    names_to      = "column",
    names_pattern = "(\\w+)-",
    values_to     = "n_not_sure"
  ) |> 
  arrange(desc(n_not_sure))
```

## Question 3

Which column in the class_survey data frame contains the greatest number of Not Sure responses?

* The package column contains the greatest number of Not Sure responses (n = 14).

# Task 10

Please recode all Not Sure responses in the class_survey data frame to missing. 

```{r}
class_survey <- class_survey |> 
  mutate(
    across(
      .cols = all_of(cols_9),
      .fns  = ~ if_else(.x == 9 | .x == 99, NA_real_, .x) 
    )
  )
```

## Question 4

Which column in the class_survey data frame contains the greatest number of missing values?

```{r}
class_survey |> 
  summarise(
    across(
      .cols = all_of(cols_9),
      .fns  = ~ sum(is.na(.x))
    )
  ) |> 
  pivot_longer(
    cols      = everything(),
    names_to  = "column",
    values_to = "n_missing"
  ) |> 
  arrange(desc(n_missing))
```

* The still_enroll column contains the greatest number of missing values (n = 21).

# Task 11

There are six variables in the class_survey data frame that have Yes and No as response options. They are required, still_enroll, asthma, insurance, hispanic, and veteran. Please create new factor versions of each of these variables. Please use the across() function to help you remove some of the repetition from your code. Please also use the “_f” variable naming convention.

```{r}
yn_cols <- c("required", "still_enroll", "asthma", "insurance", "hispanic", "veteran")
```

```{r}
class_survey <- class_survey |> 
  mutate(
    across(
      .cols  = all_of(yn_cols),
      .fns   = ~ factor(.x, 0:1, c("No", "Yes")),
      .names = "{col}_f"
    )
  )
```

## Question 5

How many people reported that they have ever been told by a doctor, nurse, or other health professional that they have asthma?

```{r}
class_survey |> 
  count(asthma_f)
```

* Nine people reported that they have ever been told by a doctor, nurse, or other health professional that they have asthma.

## Notes for students

1. Some of you may be wondering if we can use across() to apply count() to all the yes/no factor variables in one code chunk. Unfortunately, that isn't currently possible. As of this writing, we will need to use a for loop or the purrr package to repeatedly apply the count function across more than one column in our data frame.

