---
title: "Repeated Operations II"
output: html_notebook
---

# Overview

This week, we will practice using for loops and performing repeated operations with functions from the purrr package.

* [Click here to download Clients list.csv](**link update**). As you may already know, we provide short-term mental health counseling services here at the NTRHD. Once enrolled in these services, our clients are invited to come in for a maximum of six sessions. At each session, our clients complete a brief health survey. The June, July and August data was just sent to me, and I could use your help preparing it for analysis.

```{r message=FALSE}
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(readxl)
```

# Task 1

Please download ntrhd_clients_June_July_August.xlsx, import each sheet in the excel file using a for loop.

```{r}
# Save the file path to an object so we don't have to type it repeatedly 
path <- "C:/Users/jelly/OneDrive - The University of Texas Health Science Center at Houston/TA/ntrhd_clients_June_July_August.xlsx"

# Use readxl::excel_sheets to get the name of each sheet in the workbook.
# this makes our code more robust.

for(i in seq_along(sheets)) {
  # Convert sheet name to lowercase before using it to name the data frame
  new_nm <- tolower(sheets[[i]])
  assign(new_nm, read_excel(path, sheet = sheets[[i]]))
}
```

# Task 2

Now import the same file using the purrr package instead (hint: try using purrr::walk() function). 

```{r}

walk(
  .x = excel_sheets(path),
  .f = function(x) {
    new_nm <- tolower(x)
    assign(new_nm, read_excel(path, sheet = x), envir = .GlobalEnv)
  }
)
```

## Notes to students
Alternatively, you can also use the map() function to return a list of data frames and then extract each sheet into a separate data frame further (with a bit of repetition, of course!) 
## Question

1. Calculate the mean weight for each month.

```{r}
#We can implement the continuous variable calculation function from the textbook.
continuous_stats <- function(data, var) {
  data %>%  
    summarise(
      variable = quo_name(var),
      n_miss   = sum(is.na({{ var}} )),
      mean     = mean({{ var }}, na.rm = TRUE),
      median   = median({{ var }}, na.rm = TRUE),
      min      = min({{ var }}, na.rm = TRUE),
      max      = max({{ var }}, na.rm = TRUE)
    ) 
}

months<-list(june, july, august)

for (i in months){
results<-map_dfr(
  .x = quos(Height, Weight),
  .f = continuous_stats, data = i
)
print(results)}
```

* June: 66.7inches, 157lbs
July: 66.3inches, 157lbs 
August: 67.2inches, 162lbs


# Task 3. 

#Need ideas for additional tasks